---
output: github_document
---

# Hughes2009: Circadian Liver Transcriptome

[![Documentation](https://img.shields.io/badge/docs-pkgdown-blue.svg)](https://altintasali.github.io/Hughes2009/)

The `Hughes2009` package provides processed liver transcriptome data from the Hughes et al. (2009) study on circadian rhythms. It contains normalized expression data and gene-level summaries, designed for testing circadian analysis workflows and co-expression network construction.

# Installation

You can install the development version of `Hughes2009` from GitHub:

```{r install, eval=FALSE}
install.packages("pak")
pak::pak("altintasali/Hughes2009")
```

# Quick Start

This package provides an `ExpressionSet` object containing mouse liver microarray data across a 48-hour period.

```{r start, message=FALSE}
library(Hughes2009)
library(Biobase)
```

```{r data}
# Load the dataset
data(eset)

# Peek at the data
show(eset)
```

# Vignettes

```{r vig, eval=FALSE}
# View the vignette
browseVignettes("Hughes2009")
```

# Full Documentation
[Read the full documentation and tutorials online](https://altintasali.github.io/Hughes2009/)
