---
output: github_document
---

# Hughes2009: Circadian Liver Transcriptome
<!-- badges: start -->
[![Documentation](https://img.shields.io/badge/docs-pkgdown-blue.svg)](https://altintasali.github.io/Hughes2009/)
[![R-CMD-check](https://github.com/altintasali/Hughes2009/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/altintasali/Hughes2009/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The `Hughes2009` package provides processed liver transcriptome data from the Hughes et al. (2009) study on circadian rhythms. It contains normalized expression data and gene-level summaries, designed for testing circadian analysis workflows and co-expression network construction.

# Installation

You can install the development version of `Hughes2009` from GitHub:

```{r install, eval=FALSE}
install.packages("pak")
pak::pak("altintasali/Hughes2009")
```

# Quick Start

This package provides an `ExpressionSet` object containing mouse liver microarray data across a 48-hour period.

```{r start, message=FALSE}
library(Hughes2009)
library(Biobase)
```

```{r data}
# Load the dataset
data(eset)

# Peek at the data
show(eset)
```

```{r glimpse_mat}
# Extract expression matrix
expression_matrix <- exprs(eset)

# Print first 6 rows and columns
print(expression_matrix[1:6, 1:6])
```

```{r glimpse_meta}
# Extract metadata
meta <- pData(eset)

# Print first 6 rows and columns
print(meta[1:6, 1:6])
```


# Vignettes

For more details overview, please checkout the vignette: 

- [Hughes2009: Tutorial and Data Processing](https://altintasali.github.io/Hughes2009/articles/Hughes2009-tutorial.html)

# Full Documentation
[Read the full documentation and tutorials online](https://altintasali.github.io/Hughes2009/)
