---
title: "Hughes2009: Tutorial and Data Processing"
output: 
  rmarkdown::html_document:
    toc: true
    toc_float: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Hughes2009: Tutorial and Data Processing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  dpi = 300,              
  dev = "png",            
  warning = FALSE,
  message = FALSE
)

# Load necessary libraries
library(Hughes2009)
library(Biobase)
library(ggplot2)
library(limma)
```





# Installation
```{r install, eval=FALSE}
# Install via pak
if (!requireNamespace("pak", quietly = TRUE)) install.packages("pak")
pak::pak("altintasali/Hughes2009")
```

# Raw Expression Data

First, we load the raw `eset` object and perform initial quality checks to ensure data integrity.

```{r raw}
show(eset)
```
Here is how the expression matrix look like: 

```{r raw2}
expression_matrix_raw <- exprs(eset)
print(expression_matrix_raw[1:6, 1:6])
```

## Boxplots 

Visualizing expression distributions helps identify outliers or normalization issues.

```{r boxplot_1}
plot_boxplots(eset)
```

## Density Plots

Visualizing all 96 samples overlaid helps ensure there are no massive outliers or batch effects in the normalization.

```{r density_1}
plot_densities(eset)
```

## MD Plots

In the [`limma`](https://www.rdocumentation.org/packages/limma/versions/3.22.7) R package, an MD plot (Mean-Difference plot) is a specific type of MA plot used to visualize gene expression data. Its primary role in Quality Control (QC) is to identify systematic biases and assess the effectiveness of normalization. 

Below is the MD plot for the first sample: 

```{r md_1}
limma::plotMD(exprs(eset), column = 1)
```





# Filtering

Probes with low expression or inconsistent signal ([`MAS5`](https://www.rdocumentation.org/packages/affy/versions/1.44.0/topics/mas5calls) "Absent" calls) can introduce noise. We filter the dataset to retain only high-quality, reliable probes. You can play around with `min_present` parameter below to find the optimum filtering option. 

```{r filtering}
# Filter for probes present in at least 20 samples
eset_filt <- filter_mas5(eset, min_present = 20)

cat(sprintf("Original number of probes: %d\n", nrow(eset)))
cat(sprintf("Probes remaining after filtering: %d\n", nrow(eset_filt)))
```

It is also a good practice to run QC diagnostics after filtering.

## Box Plots

```{r boxplot_2}
plot_boxplots(eset_filt)
```

## Density Plots

```{r density_2}
plot_densities(eset_filt)
```

## MD Plots

```{r md_2}
limma::plotMD(exprs(eset_filt), column = 1)
```





# Collapsing by Gene Name

Since multiple probes may target the same gene, we collapse the probe-level data to the gene level using the [`WGCNA`'s](https://www.rdocumentation.org/packages/WGCNA/versions/0.94) [`collapseRows`](https://www.rdocumentation.org/packages/WGCNA/versions/0.94/topics/collapseRows) function (`MaxMean` method).

```{r collapse_rows}
eset_gene <- collapse_eset(eset_filt, method = "MaxMean")

expression_matrix_gene <- exprs(eset_gene)
print(expression_matrix_gene[1:6, 1:6])
```

## Box Plots

```{r boxplot_3}
plot_boxplots(eset_gene)
```

## Density Plots

```{r density_3}
plot_densities(eset_gene)
```

## MD Plots

```{r md_3}
limma::plotMD(exprs(eset_gene), column = 1)
```

The final `eset_gene` object is now ready for downstream analysis, such as differential expression testing with [`limma`](https://www.rdocumentation.org/packages/limma/versions/3.22.7) or gene co-expression network construction with [`WGCNA`](https://www.rdocumentation.org/packages/WGCNA/versions/0.94).