---
title: "Hughes2009: Tutorial and Data Processing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Hughes2009: Tutorial and Data Processing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  warning = FALSE,
  message = FALSE
)

# Load necessary libraries
library(Hughes2009)
library(Biobase)
library(ggplot2)
library(limma)
```

# Installation
```{r install}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("devtools")
devtools::install_github("your-username/Hughes2009")
```

# Quality Control (QC)

First, we load the raw eset object and perform initial quality checks to ensure data integrity.

## Boxplots and Density Plots

Visualizing expression distributions helps identify outliers or normalization issues.
```{r qc1}
data(eset)

# Boxplot of sample distributions
boxplot(exprs(eset), main="Sample Expression Distributions")

# Density plot
plot(density(exprs(eset)), main="Density of Normalized Data", 
     xlab="Expression Intensity", col="darkred")
```

```{r qc2}
limma::plotMD(exprs(eset),column = 1)
```

# Filtering

Probes with low expression or inconsistent signal (MAS5 "Absent" calls) can introduce noise. We filter the dataset to retain only high-quality, reliable probes.

```{r filtering}
# Filter for probes present in at least 20 samples
eset_filt <- filter_mas5(eset, min_present = 20)

cat(sprintf("Original number of probes: %d\n", nrow(eset)))
cat(sprintf("Probes remaining after filtering: %d\n", nrow(eset_filt)))
```

# Collapsing by Gene Name

Since multiple probes may target the same gene, we collapse the probe-level data to the gene level using the WGCNA collapseRows function (MaxMean method).

```{r collapse_rows}
# Collapse probes to unique gene symbols
eset_gene <- collapse_eset(eset_filt, method = "MaxMean")

# Inspect the final object
show(eset_gene)
```